### 2015-07-07 narrative/qos.tex
split into two subsections for QoS.

\paragraph{Resource Management Background:} There is an important connection between finding the resources
that can help fulfill a particular request and the performance predictability of that request, and this demands
careful resource management strategies that deal with large number of data chunks spread across storage hierarchy. 
%This connection is particularly important for data retrieval 
%- in large scale storage systems, finding a resource to read can be a significant, if not dominant, 
%component of overall read latency. 
Locating resources is a typical distributed systems problem. Centralized solutions are limited in their
scalability, because there exists a single entity that tracks updates in the system. 
Traditional file systems are good examples for this: most updates to files require updates 
to their centrally maintained block allocation tables. 
The opposite of file systems are ``pile'' systems, which, instead of keeping track of updates, search for 
resources at request time. Pile systems allow very efficient writes because clients can write whereever
there is free space and there is very low bookkeeping overhead. However, unbounded search is possible, 
if it is unclear whether the resource actually exists. 
Methods to make searches more efficient without reverting to centralized bookkeeping usually involves 
some kind of partitioning of the search space and a method to limit the search to a small number of partitions.

Frequently this involves a hash function that is relatively stable against changes to its range, such as failures or updates to the hardware of the system (see for example consistent hashing~\cite{karger:stoc97}). Consistent hashing is scalable since it can be well-known throughout a system of any size and only requires minimal movement of data in the presence of failures or system changes. Consistent hash functions can also have sets of ordered but infinite lists as their range (which works well with primary replication)~\cite{honicky:ipdps04}, implement declustering (limiting overlap between partitions), and support placement policy rules~\cite{weil:sc06}.

In this proposal we will investigate the trade-off in terms of predictable performance between ``filing'' data, making an upfront investment to ensure finding data quickly, and ``piling'' data, pushing the cost of finding to the moment when the data is actually needed. This trade-off has a parallel in data management where there are two alternative paths to achieve knowledge: the analogue to filing is ``extract, transform, and load'' (ETL) data into a warehouse, paying an upfront ingest cost to achieve fast querying; the analogue to piling is to not pay any upfront cost and query the raw data directly by using data processing platforms such as Hadoop. Recent work has found that the combination of these two techniques can be optimized to perform better than any one technique alone~\cite{lefevre:sigmod14a}. We believe that a similar opportunity exists for combining filing and piling approaches to achieve performance predictability without sacrificing too much flexibility.

@@

\paragraph{QoS Background:} 
Quality of Service (QoS) refers to the
properties of the performance quality of a particular service, in
this case the storage system.
%Performance quality can be expressed in either relative or absolute
%terms. Examples of relative performance terms are ``fair'',
%``proportional'', or ``priority/class-based'' (``the higher the
%priority the better the service'' or ``1st class is better than
%2nd''). The key advantage of relative performance terms is that
%they are easy to implement. The key disadvantage is the difficulty
%of making any guarantees based on those implementations --
%even relative guarantees are mired by the well-known effect of
%priority inversion~\cite{lampson:cacm80}. Examples of absolute
%performance terms are ``rate-based'', ``soft real-time'', and
%``hard real-time''. The key advantage of absolute performance
%terms are that they enable the implementation of strong
%guarantees: absolute terms enable agreements between application
%tasks and storage that we call \emph{reservations} and that do not
%change meaning depending
%on the storage system's workload. The key challenge of absolute
%performance terms is that their implementation is non-trivial,
%especially in large-scale systems.
One of the key differentiating features this project offers is driving storage
decisions through QoS \emph{reservations}, rather than continuing the current model
of allowing applications to compete freely without restriction, for part or all of the storage
resources. This traditional approach leads to intereference effects
\cite{lofstead:2010:io-variability,liu_hotstorage} that can greatly impact IO performance predictability. 
%Further, with the introduction of additional tiers in the
%storage stack, the performance variability will increase with greater
%competiion for limited high performance resources. 
%% Temporaily commented out the above sentence, because it may not be true for the next DOE systems. 
Unfettered, this competition can yield a lower system and application performance 
than the case where applications attempt to reserve storage resources, thereby enabling the storage system
to reason over its performance in terms of reservations and actual
performance.
In the cases where an appication arrives that needs to run more urgently
than any of the other already running applications, the system
can rapidly respond to it by reducing the amount of resources taken 
by less urgent applications, allowing the urgent application to finish,
and later reverting back to the previous committed resource level for the rest.
This can be automatically enforced via user- (or adiminstrator-) specified \textit{policies}
(section \ref{}).

>>>>

\subsubsection{Resource Management}
\paragraph{Background:} There is an important connection between finding the resources
that can fulfill a particular request and the performance predictability of that request, and this demands
careful resource management strategies that can deal with placing and retrieving a large number of data 
chunks spread across the entire storage hierarchy. 
%This connection is particularly important for data retrieval 
%- in large scale storage systems, finding a resource to read can be a significant, if not dominant, 
%component of overall read latency. 
Locating resources is a typical distributed systems problem. Centralized solutions are limited in their
scalability, because they require a single entity that tracks updates in the system. 
Traditional file systems are good examples for this - most updates to files require updates 
to their centrally maintained block allocation tables. 
The opposite of file systems are ``pile'' systems, which, instead of keeping track of updates, search for 
resources at request time. Pile systems allow very efficient writes because clients can write whereever
there is free space and there is very low bookkeeping overhead. However, unbounded search in pile systems
is possible, e.g., when it is unclear whether the resource actually exists. 
Methods to make searches more efficient without reverting to centralized bookkeeping usually involves 
partitioning of the search space and a method to limit the search to a small number of partitions.

Frequently this involves a hash function that is relatively stable against changes to its range, 
such as failures or updates to the hardware of the system. Consistent hashing~\cite{karger:stoc97}
is scalable since it can be well-known throughout a system of any size and only requires minimal 
movement of data in the presence of failures or system changes. 
%Consistent hash functions can also
%have sets of ordered but infinite lists as their range (which works well with primary replication)
%~\cite{honicky:ipdps04}, implement declustering (limiting overlap between partitions), and support
%placement policy rules~\cite{weil:sc06}.

\paragraph{Approach:} 
In this proposal we will investigate the trade-off in terms of predictable performance between 
``filing'' data, making an upfront investment to ensure finding data quickly, and ``piling'' 
data, pushing the cost of finding to the moment when the data is actually needed. This trade-off 
has a parallel in data management where there are two alternative paths to achieve knowledge: 
the analogue to filing is ``extract, transform, and load'' (ETL) data into a warehouse, paying 
an upfront ingest cost to achieve fast querying; the analogue to piling is to not pay any upfront 
cost and query the raw data directly by using data processing platforms such as Hadoop. Recent work 
has found that the combination of these two techniques can be optimized to perform better than 
any one technique alone~\cite{lefevre:sigmod14a}. We believe that a similar opportunity exists 
for combining filing and piling approaches to achieve performance predictability without sacrificing 
too much flexibility.

@@

\subsubsection{QoS}
\paragraph{Background:} Quality of Service (QoS) refers to the
properties of the performance quality of a particular service, in
this case the storage system.
%Performance quality can be expressed in either relative or absolute
%terms. Examples of relative performance terms are ``fair'',
%``proportional'', or ``priority/class-based'' (``the higher the
%priority the better the service'' or ``1st class is better than
%2nd''). The key advantage of relative performance terms is that
%they are easy to implement. The key disadvantage is the difficulty
%of making any guarantees based on those implementations --
%even relative guarantees are mired by the well-known effect of
%priority inversion~\cite{lampson:cacm80}. Examples of absolute
%performance terms are ``rate-based'', ``soft real-time'', and
%``hard real-time''. The key advantage of absolute performance
%terms are that they enable the implementation of strong
%guarantees: absolute terms enable agreements between application
%tasks and storage that we call \emph{reservations} and that do not
%change meaning depending
%on the storage system's workload. The key challenge of absolute
%performance terms is that their implementation is non-trivial,
%especially in large-scale systems.
One of the key differentiating features this project offers is driving storage
decisions through QoS \emph{reservations}, rather than continuing the current model
of allowing applications to compete freely without restriction, for part or all of the storage
resources. This traditional approach leads to intereference effects
\cite{lofstead:2010:io-variability,liu_hotstorage} that can greatly impact IO performance predictability. 
%Further, with the introduction of additional tiers in the
%storage stack, the performance variability will increase with greater
%competiion for limited high performance resources. 
%% Temporaily commented out the above sentence, because it may not be true for the next DOE systems. 
Unfettered, this competition can yield a lower system and application performance 
than the case where applications attempt to reserve storage resources, thereby enabling the storage system
to reason over its performance in terms of reservations and actual
performance.
In the cases where an appication arrives that needs to run more urgently
than any of the other already running applications, the system
can rapidly respond to it by reducing the amount of resources taken 
by less urgent applications, allowing the urgent application to finish,
and later reverting back to the previous committed resource level for the rest.
This can be automatically enforced via user- (or adiminstrator-) specified \textit{policies}
(section \ref{}).
